A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-08 podc_paper/correctness-2.tex
working on GC section

We next prove the analogue of Lemma \ref{lem::headProgress}.
\begin{customlemma}{$4'$}\label{lem::headProgressGC}
 If $h>0$ and a block with index $h$ has been inserted into $v.\fld{blocks}$ then 
 $v.\fld{blocks}[h-1].\fld{end\sub{left}} \leq v.\fld{blocks}[h].\fld{end\sub{left}}$ and 
 $v.\fld{blocks}[h-1].\fld{end\sub{right}} \leq v.\fld{blocks}[h].\fld{end\sub{right}}$.
\end{customlemma}
\begin{proof}
The block $B$ with index $h$ was installed into $v$'s RBT by the CAS at line \ref{casGC} 
that changed the tree from $T$ to $T'$.
Before this CAS, line \ref{refreshReadTGC} read the tree $T$ from $v.blocks$,
line \ref{refreshReadMax} found a block $B'$  with \fld{index} $h-1$ in $T$,
and then line \ref{invokeCreateBlockGC} created the block $B$ with $\fld{index}=h$.
Since $B'$ was already in $T$ before $B$ was created, the
\op{CreateBlock}$(v,b-1)$ that created $B'$ terminated before the 
\op{CreateBlock}$(v,b)$ that created $B$ started.
By \Cref{nonEmptyIncreasing}, the value that line \ref{createEndLeftGC} of \op{CreateBlock}$(v,b-1)$ 
stores in $B'.\eleft$ is less than or equal to the value that line \ref{createEndLeftGC} 
of \op{CreateBlock}$(v,b)$ stores in $B.\eleft$.  
Similarly, the values stored in $B'.\eright$ and $B.\eright$ at line \ref{createEndRightGC} satisfy the claim.
\end{proof}

@@

\begin{customlemma}{$9'$}
\label{successfulRefreshGC}
Let $R$ be a call to \op{Refresh}($v$) that performs a successful \op{CAS} on line \ref{casGC} (or terminates at line \ref{addOPGC}).
In the configuration after that CAS (or termination, respectively), $v$ contains all operations that $v$'s children contained 
when $R$ executed line~\ref{refreshReadTGC}.
\end{customlemma}
\begin{proof}
Suppose $v$'s child (without loss of generality, $v.\fld{left}$) contained an operation $op$ 
when $R$ executed line \ref{refreshReadTGC}.
Let $i$ be the index of the block containing $op$ that was in $v$'s RBT before 
$R$ executed line \ref{refreshReadTGC}.
We consider two cases.

@@

\begin{lemma}\label{lem::doubleRefreshGC}
Consider two consecutive terminating calls $R_1$, $R_2$ to \op{GCRefresh}($v$) by the same process.
All operations propagated to $v$'s children when $R_1$ begins
are propagated to $v$ when $R_2$ terminates.
\end{lemma}
\begin{proof}
\end{proof}

@@

>>>>

We next prove the analogue of Lemma \ref{lem::headProgress}.
\begin{customlemma}{\ref{lem::headProgress}$'$}\label{lem::headProgressGC}
 If $h>0$ and a block with index $h$ has been inserted into $v.\fld{blocks}$ then 
 $v.\fld{blocks}[h-1].\fld{end\sub{left}} \leq v.\fld{blocks}[h].\fld{end\sub{left}}$ and 
 $v.\fld{blocks}[h-1].\fld{end\sub{right}} \leq v.\fld{blocks}[h].\fld{end\sub{right}}$.
\end{customlemma}
\begin{proof}
The block $B$ with index $h$ was installed into $v$'s RBT by the CAS at line \ref{casGC} 
that changed the tree from $T$ to $T'$.
Before this CAS, line \ref{refreshReadTGC} read the tree $T$ from $v.blocks$,
line \ref{refreshReadMax} found a block $B'$  with \fld{index} $h-1$ in $T$,
and then line \ref{invokeCreateBlockGC} created the block $B$ with $\fld{index}=h$.
Since $B'$ was already in $T$ before $B$ was created, the
\op{CreateBlock}$(v,b-1)$ that created $B'$ terminated before the 
\op{CreateBlock}$(v,b)$ that created $B$ started.
By \Cref{nonEmptyIncreasing}, the value that line \ref{createEndLeftGC} of \op{CreateBlock}$(v,b-1)$ 
stores in $B'.\eleft$ is less than or equal to the value that line \ref{createEndLeftGC} 
of \op{CreateBlock}$(v,b)$ stores in $B.\eleft$.  
Similarly, the values stored in $B'.\eright$ and $B.\eright$ at line \ref{createEndRightGC} satisfy the claim.
\end{proof}

@@

\begin{customlemma}{\ref{successfulRefresh}$'$}
\label{successfulRefreshGC}
Let $R$ be a call to \op{Refresh}($v$) that performs a successful \op{CAS} on line \ref{casGC} (or terminates at line \ref{addOPGC}).
In the configuration after that CAS (or termination, respectively), $v$ contains all operations that $v$'s children contained 
when $R$ executed line~\ref{refreshReadTGC}.
\end{customlemma}
\begin{proof}
Suppose $v$'s child (without loss of generality, $v.\fld{left}$) contained an operation $op$ 
when $R$ executed line \ref{refreshReadTGC}.
Let $i$ be the index of the block containing $op$ that was in $v$'s RBT before 
$R$ executed line \ref{refreshReadTGC}.
We consider two cases.

@@

\begin{customlemma}{\ref{lem::doubleRefresh}$'$}\label{lem::doubleRefreshGC}
Consider two consecutive terminating calls $R_1$, $R_2$ to \op{Refresh}($v$) by the same process.
All operations contained $v$'s children when $R_1$ begins
are contained in $v$ when $R_2$ terminates.
\end{customlemma}
\begin{proof}
If either $R_1$ or $R_2$ performs a successful \op{CAS} at line \ref{casGC} or terminates at line \ref{addOPGC}, the claim follows
from \Cref{successfulRefreshGC}.
So suppose both $R_1$ and $R_2$ perform a failed \op{CAS} at line \ref{casGC}.
\here{To be continued....}

\end{proof}

\here{next, should probably show that blocks are removed only after they are no longer needed; i.e., every dequeue has its answer (either returned, or written in its response field) and every enqueue has a matching dequeue that has either terminated or has a response written.}

@@

\subsection{Analysis}

We first bound the size of RBTs.  Let $q_{max}$ be the maximum size of the queue at any time during the sequential execution given by the linearization $L$.

\begin{lemma}
If the maximum \fld{index} in a node's RBT is a multiple of $p^2$, then
it contains $O(q_{max}+p)$ blocks.
\end{lemma}

\begin{corollary}\label{RBTbound}
At all times, the size of a node's RBT is $O(q_{max}+p^2)$. 
\end{corollary}
\begin{proof}

\end{proof}

\begin{proposition}
The amortized step complexity is $O(\log p \log(p+q_{max}))$ per operation.
\end{proposition}
\begin{proof}
It follows from \Cref{RBTbound}
that all the RBT routines we use to perform \op{Split}, \op{Insert} and searches for
blocks with a particular index or for a \fld{sum\sub{enq}} value (in line \ref{FRsearchGC} or \ref{getChildGC}) can all be done in $O(\log(p+q_{max}))$ steps.

First, we bound the number of steps taken \emph{excluding} the GC phase in line \ref{GCstart}--\ref{GCend}.
An \op{Enqueue} or null \op{Dequeue} does $O(1)$ RBT operations and other work at each level of the tree during \op{Propagate},
for a total of $O(\log p \log(p+q_{max}))$ steps.
A non-null \op{Dequeue} must also search for a block in the root at line \ref{FRsearchGC}
and call \op{GetEnqueue}, which does $O(1)$ RBT operations (including a search at line \ref{getChildGC}) and other work at each level of the tree.
Thus, a \op{Dequeue} also takes $O(\log p \log(p+q_{max}))$ steps.

Now we consider the additional steps a process takes to do GC in line \ref{GCstart}--\ref{GCend}.
\here{To be continued}

\end{proof}
